Results_setup

### 2020-04-08 src/Result_Setup.Rmd
generating file for organizing DF and other results

---
title: "MLH1 Results Outline"
date: "null"
output:
  word_document:
    toc: yes
    toc_depth: '5'
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
bibliography: refs_Main_Report.bib
---

@@

<!-- 
two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
-->

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1

INSERT TABLE FOR THE TOTAL SC means ect

The difference between

bin is short for binary, it reports the binarized area of the SC

skel -- takes the skeletonized midline of the bin objects
-->

```{r table.for.plot, echo=FALSE, warning=FALSE}
#why do I add in a different table?

#this move this to the setup file
#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed

#pooled
strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)

                       #range_mouse_means = range()
                   #consider adding crhm proportions?
    )

strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")

#adjust order of plot here
strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ",
  "WSB","G","LEW","PERC",
                     "CAST",
                         "SPRET","SPIC" ), order=T )

```{r Fig1, echo=FALSE, warning=FALSE}

#might need to nudge the X values

Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#remove y labfor male

   #segments
#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
 # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label

 #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
   #        y = c(18.5,18.5,33,18.5, 18.5),

  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
    #       fontface = 'italic', hjust = 0.5, size=3.5)

```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")

#table( droplevels(female_mouse.avs_for.analysis$strain) )
#there are cast strains too

female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))

#female.specific.plot

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
#try customize the plot size

male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)

#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")

just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)

#table( droplevels(just.male_mouse.av$strain) )
#there are cast strains too

Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

#fomo.male_mouse.av

male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
male.plot

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))

```{r three.plots, echo=FALSE, warning=FALSE}
#merge the 3 plots
#try patchwork

#Fig1_v1_strain + female.specific.plot + male.plot

#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
#Fig1_v1_strain / (female.specific.plot + male.plot )

#(p1 | p2 | p3)/p4
```

```{r main.fig1s, echo=FALSE, include=FALSE}
strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )

Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point( size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

#strain.table.HQ_Fig.og

strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")

strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
  "WSB","G","LEW"), order=T )

Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

```{r basic.numbers, echo=FALSE, eval=FALSE}

#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )

#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )

```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
#str(AP_mouse_table_w.Ages)

AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

#remove sex.x == na, and strain.x=na or F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
#plot variance and mean (facet sex)

AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]

VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")

cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")

#str(AP_mouse_table.HQ)

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
#[1] 0.5750399
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
#0.321601

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
# 0.3515544
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
#-0.03713536

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
# p-value = 8.623e-09

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )

#remove mice with fewer than 10 cells

AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]

VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")

cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])

cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )

#why did I male just a female table?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]

female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)
```

# genome-wide recombination rate estimates for both sexes

```{r MLH1.summ.stats, echo=FALSE}

#strain.table.HQ_Fig.og  all strains
#strain.table.HQ_Fig  extra strains removed

main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")

main.table <- add_subsp(main.table)
main.table <- add_species(main.table)

#main.table %>% filter(sex == "female")
#main.table %>% filter(sex == "male")

#G.HETC.val #1.2
#LEW.HETC.val  #1
#WSB.HETC.val  #1
#KAZ.HETC.val  #1.1

#SKIVE.HETC.val # 0.96
#PWD.HETC.val  #0.88
#MSM.HETC.val #0.9

#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")

#strain.table.HQ_Fig

#sum( (main.table  %>% filter(sex == "male"))$Ncells)
#the kable version is bulkier than the other versaion

```{r MLH1.sum.stats.Bret.edits, echo=FALSE}

AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
#total mice
#length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )

#average cell number for spermatocytes and oocytes
#mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
#mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
```

We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M. m. domesticus*, *M. m. musculus* and *M. m. molossinus*. Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )` mice were quantified from an average of `r mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes).

Graphical comparisons between the two sexes reveal several patterns (Figure 1A). First, recombination rate differs between females and males in most strains. Second, the direction and magnitude of heterochiasmy varies among strains. Although the majority of strains show higher recombination rates in females (following the pattern in laboratory mice), two musculus strains and one molossinus strain exhibit male-biased heterochiasmy. Hence, relative recombination rates in the two sexes are evolutionarily labile.

The general patterns of heterochiasmy in house mouse are displayed in Figure 1A. Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_. Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.

Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C). Female recombination rates are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure 1B). In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).

```{r female.strain.table, echo=FALSE}
female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#F.strain.table

#F.strain.table

# I don't know why most of the table is blank
```

```{r female.dif, echo=FALSE}

#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)
```

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

#variance plots
musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.female.VAR.plot

#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
#Mol.female.VAR.plot

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.female.VAR.plot

```{r Female.specific.glm, echo=FALSE}

glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)

summry.female.strain.effects <- summary(glm.female.strain.subsp)
#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)

MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)

#round(lm.female.strain_sum$coefficients[,c(1,4)],5)

#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
```

```{r FEMALE.VAR.MODELS, echo=FALSE}
## within mouse variance models 
lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)

lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
lmVAR.female.strain_sum <- summary(lmVAR.female.strain)

#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]
#Q12
Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")

lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)

lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD

lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)

#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
#LEW sig

```
<!-- male section -->

```{r male.strain.table, echo=FALSE}
M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        totNcells  = sum(Ncells, na.rm=TRUE),

                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm=TRUE),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#M.strain.table
#same issue as the female table .. ugh
```

```{r male.diff, echo=FALSE}
just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)

#how much greater are the High Rec gwRRs
just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))

#how much bigger are the high rec male means?
sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
#mean(sans_LewRec_male$mean_co)

PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)

MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)

SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
#mean(SKIVE_male$mean_co) /  
```

```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}

#variance plots
musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)

```{r Male.specific.glm, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect
glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)

#summry.male.strain.effects <- summary(glm.male.strain.subsp)
#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)

#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)

#1. determine which strain effects are sig

#M2
MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)

#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)

#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)

#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

```{r MALE.SPECIFIC.variance.model, echo=FALSE}
## within mouse variance models 
lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)

#all p values kinda high, MSM and TOM are slightly sig
lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
#non sig

lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
lmcv.male.strain_sum <- summary(lmcv.male.strain)
#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant

#Q12
Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")

lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
#NS

lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
#NS

# Partitioning variation in recombination rate

```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}

#DF.HetC.MixedModel.HQ

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1.MM.M1)$coefficients

MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)

```{r MLH1.glmM2, echo=FALSE}

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed
MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )

summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)

#table of just coeff anf pvalue
#round(summ$coefficients[,c(1,4)],5 )

#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
```

```{r MLH1.glm.M3, echo=FALSE}

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])

summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)

#round(summ.strain.fixed$coefficients[,c(1,4)],5)

#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
```

To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice. We began with a full mixed model (M1, see Methods), which showed that strain 
(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results` -- 3.9710^{-5}), subspecies (p=`r MLH1.MMlmer_results$sub.sp`--  1.7210^{-4}), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval` -- 3.110^{-5}) each significantly affect recombination rate.

After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex. A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p= 3.9910^{-6} p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
and domesticus^G^ (p= 1.0410^{-6}  -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`). In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).

>>>>

---
title: "Results_setup"
author: "April Peterson"
date: "April 7, 2020"
output: html_document
---

@@

# SAVE the data
